\PassOptionsToPackage{prologue,dvipsnames}{xcolor}
\documentclass[twocolumn]{aastex631}

\usepackage{showyourwork}
\usepackage{bm}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{appendix}
\usepackage{ifthen}
\usepackage{array}
\usepackage{soul}
\usepackage{float}
\input{macros.tex}
\usepackage{multirow}
\newcommand{\CIPlusMinus}[1]{{#1[median]^{+#1[error plus]}_{-#1[error minus]}}}
\newcommand{\CIPlusMinusPer}[1]{{#1[median]\%^{+#1[error plus]\%}_{-#1[error minus]\%}}}
\newcommand{\CIBoundsBracket}[1]{{[#1[5th percentile], #1[95th percentile]]}}
\newcommand{\CIBoundsDash}[1]{{#1[5th percentile]\textendash#1[95th percentile]}}

\newcommand{\msun}{\ensuremath{{\,\rm M}_\odot}}
\newcommand{\popA}{\textsc{SpinPop\textsubscript{A}}}
\newcommand{\popB}{\textsc{SpinPop\textsubscript{B}}}
\newcommand{\first}{\popA{}\textsc{:Peak}}
\newcommand{\contB}{\popB{}\textsc{:Continuum}}
\newcommand{\contA}{\popA{}\textsc{:Continuum}}
\newcommand{\brucepaper}{\citet{2022arXiv221012834E}}
\newcommand{\othreea}{\citet{2021ApJ...913L...7A}}
\newcommand{\cont}[1]{\textcolor{red}{DELTE THIS}}
\newcommand{\second}[1]{\textcolor{red}{DELTE THIS}}
\newcommand{\base}{\textsc{Isolated Peak Model}}
\newcommand{\comp}{\textsc{Peak+Continuum Model}}


\newcommand{\result}[1]{\textcolor{BurntOrange}{#1}}
\newcommand{\bruce}[1]{\textcolor{blue}{BE: #1}}
\newcommand{\jaxen}[1]{\textcolor{Cerulean}{JG: #1}}
\newcommand{\ben}[1]{\textcolor{orange}{BF: #1}}
\newcommand{\NewChange}[1]{\textcolor{BrickRed}{#1}}
\newcommand{\remove}[1]{\textcolor{WildStrawberry}{REMOVE: #1}}




\graphicspath{{./}{figures/}}

\begin{document}
\title{Cosmic Cousins: Identification of a Subpopulation of Binary Black Holes Consistent with Isolated Binary Evolution}

\author{Jaxen Godfrey}
\email{jaxeng@uoregon.edu}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
\author{Bruce Edelman}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}
\author{Ben Farr}
\affiliation{Institute  for  Fundamental  Science, Department of Physics, University of Oregon, Eugene, OR 97403, USA}

% Abstract with filler text
\begin{abstract}
    Observations of gravitational waves (GWs) from merging compact binaries have become a regular occurrence. The continued advancement of the LIGO-Virgo-KAGRA (LVK) detectors have now produced a catalog of over 90 such mergers, from which we can begin to uncover the formation history of merging compact binaries. In this work, we search for subpopulations in the LVK's third gravitational wave transient catalog through the use of data-driven mixture models in a hierarchal Bayesian inference framework. By allowing for unique correlations between mass and spin in each subpopulation, we find an over density of mergers with a primary mass of $\result{\CIPlusMinus{\macros[Mass][Base][PeakA][max]}\msun}$, consistent with isolated binary formation \NewChange{and stable mass transfer}. This low-mass subpopulation has a spin magnitude distribution peaking at $a_\mathrm{peak}=$ \result{$\CIPlusMinus{\macros[SpinMag][Base][PeakA][max]}$}, exhibits spins preferentially aligned with the binary's orbital angular momentum, is constrained by \result{$\CIPlusMinus{\macros[NumEvents][Base][PeakA]}$} of our observations, and contributes \result{$\CIPlusMinusPer{\macros[BranchingRatios][Base][PeakA][Percent]}$} to the overall population of BBHs. We find a $\sim\macros[FracCut][BgreaterA]\%$ chance that the bulk of the events in the $15-55\msun$ range share a spin distribution with the $10\msun$ peak, with $99\%$ of these events possessing primary masses less than $m_{1,99\%} = $ \result{$\CIPlusMinus{\macros[Mass][Composite][ContinuumA][99percentile]}\msun$}, providing an estimate of the lower edge of the theorized pair instability mass gap. Additionally, we find mild evidence for a subpopulation of high mass BBHs near $60\msun$. This work is a first step in gaining a deeper understanding of compact binary formation and evolution with data-driven models, and will provide more robust conclusions as the catalog of observations becomes larger. 
\end{abstract}

\input{intro.tex}
\input{methods.tex}
\input{results.tex}
\input{conclusion.tex}

\section{Acknowledgements}\label{sec:acknowledments}
This material is based upon work supported by the National Science Foundation Graduate Research Fellowship under Grant No. 2236419. This research has made use of data, software and/or web tools obtained from the Gravitational Wave Open Science Center 
(\url{https://www.gw-openscience.org/}), a service of LIGO Laboratory, the LIGO Scientific Collaboration and the Virgo Collaboration. 
The authors are grateful for computational resources provided by the LIGO Laboratory and supported by National Science Foundation Grants PHY-0757058 and PHY-0823459.  
This work benefited from access to the University of Oregon high performance computer, Talapas. This material is based upon work supported 
in part by the National Science Foundation under Grant PHY-2146528 and work supported by NSF's LIGO Laboratory which is a major facility 
fully funded by the National Science Foundation.
\software{
\textsc{Showyourwork}~\citep{2110.06271},
\textsc{Astropy}~\citep{1307.6212, 1801.02634, 2206.14220},
\textsc{NumPy}~\citep{10.1038/s41586-020-2649-2},
\textsc{SciPy}~\citep{10.1038/s41592-019-0686-2},
\textsc{Matplotlib}~\citep{10.1109/MCSE.2007.55},
\textsc{Jax}~\citep{github.com/google/jax},
\textsc{NumPyro}~\citep{1810.09538,1912.11554},
}
\bibliography{bib}{}
\bibliographystyle{aasjournal}

% \variable{output/event_cat_table.tex}

\input{appendices.tex}

\end{document}